---
title: "Initial Korean covid 19"
author: "leia"
date: "fall 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
 
 
```{r}
str(kRdata)
```
  
# Your Question - Is dying independent of sex?

$H_0$ there is no difference in dying based on sex

$H_A$ there is no difference in dying bases on sex 

## Method

the variable sex is categorical and the variable results are categorical, so therefore we us cat~cat

## Results 

## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
this bar chart shows the percentage of male and female who died from covid, and shows people who had it but are still alive. 

## Numerical results 

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

this table shows the percentageof male and female who had covid that is alive and the percentage of who died. male survivors percentage is 44% and the female surviors is 56%. 

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

this table show the p value in the chi-square stat, which is 0.0012. the p calue in the fisher test, which is 0.0009. therefore we reject both null hypothesis.  


## Conclusions

